---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# suitestrings

<!-- badges: start -->

[![Repo status: work in progress](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)

<!-- badges: end -->

This is primarily a personal learning project, but `{suitestrings}` aims to provide comprehensive and convenient set of functions for working with strings in R. More specifically: 

- Combining strings.
- Cleaning and transforming strings.
- Manipulating strings based on regular expression patterns.

A few benefits of the design of this package:

-   A consistent naming scheme - Including prefixes designed for use with auto-complete:

    -   `str_` for vectorised operations on strings.

    -   `chr_` for vector-wide operations over strings.

-   Order of inputs - The string or strings to work on is usually used as the first argument instead of the pattern, which makes it easier to use with pipes than base R.

-   No additional dependencies - this package has no Imports so you'll only need base R to use it.

## Installation

You can install suitestrings from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("davidfoord1/suitestrings")
```

## Examples
```{r}
library(suitestrings)
```
### Combine strings together

Concatenation across multiple vectors.

```{r}
str_concat(c("Mary ", "Joseph "), 
           "had a ", 
           c("little ", "technicolour "), 
           c("lamb ", "dreamcoat "))
```

Evaluate R expression in strings

```{r}
x <- 25
str_glue("{x} squared is {x^2} and the square root of {x} are {sqrt(x)}")
```

### Clean and transform strings

```{r}
str_to_snake_case("  This IS  a->>GreAt<<-strING!!")
```

### Manipulating strings based on regular expression patterns

```{r}
# Prepare a vector of strings
strings <- c("flat-hat", "backpack", "roll", "cat-sat-on-a-mat")
# Define a pattern for any three characters with "a" in the middle
regex_pattern <- "\\wa\\w"
```

Detect, extract and replace patterns in strings

```{r}
str_detect(strings, regex_pattern)
str_extract_first(strings, regex_pattern)
str_replace_first(strings, regex_pattern, "lea")
```

Work with specific occurrences of patterns

```{r}
# Get the second match for each string
str_extract_nth(strings, regex_pattern, 2)

# Replace the first match in each string
str_remove_last(strings, regex_pattern)
```

Work with every occurrence of a pattern
```{r}
# Get a list of every match for each string
str_extract_all(strings, regex_pattern)

# Get every match from `strings` into one character vector
chr_extract_all(strings, regex_pattern)

# Get the subset of `strings` that contain a match
chr_subset(strings, regex_pattern)
```
