---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(suitestrings)
```

# suitestrings

<!-- badges: start -->

<!-- badges: end -->

This is primarily a learning project, for me to build my understanding of string manipulation and regular expressions (both using base R and popular packages) as well as how to create packages.

All `{suitestrings}` functionality is implemented in base R; most functions are built as wrappers around `` `{base}` `` string manipulation functions. The interface largely mimics what's seen in [stringr](https://stringr.tidyverse.org/reference/index.html) and [stringi](https://stringi.gagolewski.com/rapi/stri_count.html), with the intent of being expressive and easily understandable in English. As such, it is set up to be used *instead* of those packages.

So, the goal of suitestrings is to provide comprehensive set of functions for string manipulation. Some notable features:

-   Consistent naming scheme - Including prefixes designed for use with autocomplete :

    -   `str_` for vectorised operations on strings.

    -   `chr_` for vector-wide operations over strings.

-   Order of inputs - The string(s) to work on is usually used as the first argument instead of the pattern, which makes it easier to use with pipes than base R.

-   No additional dependencies - this package has no Imports and doesn't do anything you can't replicate exactly in base R.

## Installation

You can install suitestrings from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("davidfoord1/suitestrings")
```

## Examples

``` {r}
# Prepare a vector of strings
strings <- c("flat-hat", "backpack", "roll", "cat-sat-on-a-mat")
# Define a pattern for a three character string with a in the middle
regex_pattern <- "\\wa\\w"

# Get the second match from each string
str_extract_nth(strings, regex_pattern, 2)
```
