---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(suitestrings)
```

# suitestrings

<!-- badges: start -->

[![Repo status: work in progress](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)
[![R-CMD-check](https://github.com/davidfoord1/suitestrings/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/davidfoord1/suitestrings/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

This is primarily a personal learning project. It aims to provide a comprehensive and convenient set of functions for working with strings in R. More specifically, for:

-   Combining strings.
-   Cleaning and transforming strings.
-   Manipulating strings based on regular expression patterns.

The interface is built with a consistent naming scheme and argument structure. The first argument will be the string or strings to work on, which makes it particularly convenient to work with pipes. Functions start with prefixes for easy identification and auto-completion:

-   `str_` for vectorised operations on strings.
-   `chr_` for vector-wide operations over strings.

You can read more at `vignette("suitestrings-conventions")`.

## Installation

You can install suitestrings from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("davidfoord1/suitestrings", build_vignettes = TRUE)

# Load the package to your library
library(suitestrings)
```

This package is built with base R, so there are no additional dependencies to install.

## Examples

### Combine strings together

#### Concatenate strings

```{r}
str_concat("Mary", "had", "a", "little", "lamb", separator = " ")
```

#### Evaluate R expression in strings

`str_glue()` treats text in braces `{}` like R code

```{r}
statement <- "{x} squared is {x^2} and its square root is {sqrt(x)}."

x <- 25
str_glue(statement)

x <- 16
str_glue(statement)
```

### Clean and transform strings

```{r}
str_to_snake_case("  This /IS/  a ->>!!GREAT!!<<-STRing!!")
```

### Manipulate strings based on regular expression patterns

```{r}
# Prepare an example character vector.
strings <- c("flat-hat", "backpack", "roll", "cat-sat-on-a-mat")
# Define a pattern for a three letter sequence with "a" in the middle.
pattern <- "\\wa\\w"
```

#### Detect, extract and replace patterns in strings

```{r}
# Does a string contain a match?
str_detect(strings, pattern)
# Get the first match in a string
str_extract_first(strings, pattern)
# Replace the first match in a string
str_replace_first(strings, pattern, "rat")
```

#### Work with specific occurrences of patterns

Using suffixes _first, _nth and _last

```{r}
# Get the second match for each string
str_extract_nth(strings, pattern, 2)
# Remove the last match in each string
str_remove_last(strings, pattern)
```

#### Work with every occurrence of a pattern

```{r}
str_locate_all(strings, pattern)

# Get a list of every match for each string
str_extract_all(strings, pattern)
```

#### Work on the character vector as a whole

```{r}
# Get every match from `strings` into one character vector
chr_extract_all(strings, pattern)

# Which elements of the vector contain a match?
chr_which(strings, pattern)

# Get the subset of `strings` that contain a match
chr_subset(strings, pattern)
```
