---
title: "suitestrings"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{suitestrings}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(suitestrings)
```

# Introduction

## Character Vector Basics

This package is for operating on three key kinds of things.

1.  **Characters** - specifically, individual unicode code points.

```{r}
"A"
"\u42"
# This includes non alpha-numeric characters, like whitespace and punctuation
"\u020"
```

2.  **Strings** - sequences of characters.

```{r}
"Hello"

# How many characters are in the string?
str_length("\u42")
str_length("\u020")
str_length("Hello")
```

3.  **Character vectors** - sequences of strings.

```{r}
c("Hello", "world")

# What is the object type?
typeof(c("Hello", "world"))

# How many elements are in the vector?
length(c("Hello", "world"))
```

Our function names use a shorthand `char`, for characters, `str` for strings and `chr` for character vectors, with the significance of the distinctions detailed more in `vignette(suitestrings-conventions)`.

It's useful to keep in mind that for many purposes the three are all the same; as R objects, they are all stored as character vectors i.e. a single string is simply a character vector of `length()` 1. Usually when we say something operates on a string, it will work on every individual string within a character vector.

```{r}
# Viewed for their vector properties,
# a character and a string look much the same
typeof("A")
typeof("Hello")

length("A")
length("Hello")
```

## suitestrings operations

We offer a few families of operations for working with strings:

1.  Character-based transformation: Change strings based on individual characters within them. Extend strings by adding characters, shorten strings by removing them. Handle specific kinds of characters like whitespace.
2.  String combination: Concatenate strings together, with other strings or even with the results of R expressions converted to characters,
3.  Pattern matching operations: Manipulate strings based on a pattern of characters, which is usually defined by a regular expression.
4.  Character vector organisation: Sort a character vector alphabetically or remove duplicate elements of a character vector.

#### Character-based transformation

There are some functions useful for quickly cleaning a string. `str_trim()` removes whitespace from the end and `str_squish()` additionally reduces whitespace in the middle to a single character:

```{r}
str_squish("  Get    these    spaces \u020         under       control    ")
```

The somewhat opposite functions `str_pad()` and `str_indent()` fill a string to a minimum length, and add a specific number of spaces, respectively.

```{r}
str_pad("hello", 10)
str_indent("hello", 3)

# Though they can both extend with other characters
str_indent("hello", 3, indent = ".")
```

Functions to change case can help with consistent formatting:

```{r}
str_to_upper_case("hello")

str_to_snake_case("hEllO  woRLd")
```

#### String combination

Concatenate strings together with `str_concat()` or `str_glue()`.

```{r}
str_concat("abc", "def") 
# Both can take a custom separator argument to place betwen strings
str_glue("abc", "def", separator = " ")
```

They can also combine strings with R expressions (coerced to `as.character()`). `str_glue()` is designed to handle this more elegantly by treating text in curly braces `{}` as R code.

```{r}
str_concat(
  "hello", 
  10 * 10,
  "worlds",
  separator = " "
)

str_glue("hello {10*10} worlds")
```

Concatenate the elements of a character vector into a single string with `chr_collapse()`

```{r}
chr_collapse(c("abc", "def"))
chr_collapse(1:5, separator = ", ")
```

#### Pattern matching operations

We offer functions to **detect**, **locate**, **extract**, **replace**, **remove** and **split** patterns of characters in strings. Except for the detect family, suffixes `_first()`, `_nth()` and `last()` are used to specify pattern occurences within strings, and \_`all()` to work with every occurrences.

A pattern by default is a regular expression, a type of string with a special set of characters defined that allow it to represent many different strings. If you wish to use a string literally as a position you can supply argument `fixed = TRUE`.

#### Character vector organisation
