---
title: "suitestrings"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{suitestrings}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(suitestrings)
```

## Character vector basics

This package is for operating on three key kinds of things:

### Strings

Sequences of characters, delimited by either double (`"`) or single quotes (`'`).

```{r}
"Hello"
'Everyone'
```

### Characters

Any letter or number like `A` or `1`, but more specifically, any individual unicode code point:

```{r}
"A"

# Whitespace and punctuation are characters too
" "

# Or even emojis...
"😮"

# There are other ways to represent a single character
"\u42"
```

Of course, those were *all* strings.

```{r}
# How many characters are in each of these strings?
str_length("Hello")
str_length('Everyone')
str_length("A")
str_length(" ")
str_length("😮")
str_length("\u42")
```

### Character vectors

Sequences of strings.

```{r}
c("Hello", "world")

# What is the object type?
typeof(c("Hello", "world"))

# How many elements are in the vector?
length(c("Hello", "world"))
```

Our function names use a shorthand `str` for strings and `chr` for character vectors, with the significance of the distinctions detailed more in `vignette(suitestrings-conventions)`.

It's useful to keep in mind that for many purposes the three are all the same; as R objects, they are all stored as character vectors i.e. a string is simply a character vector with a single element, which may contain no, one or many characters. Usually when we say something operates on a string, it will work on every individual string within a character vector.

```{r}
# Viewed for their vector properties,
# a character and a string look much the same:

# A character vector
typeof("A")
typeof("Hello")

# With one element
length("A")
length("Hello")
```

## suitestrings operations

There are a few families of operations for working with strings:

1.  Character-based transformation: Change strings based on individual characters within them. Extend strings by adding characters, shorten strings by removing them. Handle specific kinds of characters like whitespace.
2.  String combination: Concatenate strings together, with other strings or even with the results of R expressions converted to characters,
3.  Pattern matching operations: Manipulate strings based on a pattern of characters, which is usually defined by a regular expression.
4.  Character vector organisation: Sort a character vector alphabetically or remove duplicate elements of a character vector.

### Character-based transformation

#### Shortern strings

There are some functions useful for quickly cleaning a string. `str_trim()` removes whitespace from the ends and `str_squish()` additionally reduces whitespace in the middle to a single character:

```{r}
str_trim("  Get rid of spaces at the ends          ")
str_squish("  Get    these    spaces \u020         under       control    ")
```

If you want cut a string down to size, but also to have the string indicate that it has been shortened, you can **truncate** it down to a specified number of characters:

```{r}
str_truncate("Sometimes we just need to make a string smaller", 20)

# To cut it off without an ellipsis
str_truncate("Sometimes we just need to make a string smaller", 20, ellipsis = "")
```

#### Extend strings
The somewhat opposite functions `str_pad()` and `str_indent()` fill a string to a minimum length, and add a specific number of spaces, respectively.

```{r}
str_pad("hello", 10)

# Though they can both extend with other characters
str_indent("hello", 3, indent = ".")
```

#### Convert Case
Functions to change case can help with consistent formatting:

```{r}
str_to_upper_case("hello")

str_to_snake_case(c("nO  FUNny  buSSIneSs", "    TIDY  this  UP  "))
```

### String combination

Concatenate strings together with `str_concat()` or `str_glue()`.

```{r}
str_concat("abc", "def") 

# Both can take a custom separator argument to place between strings
str_glue("abc", "def", separator = " ")
```

They can also combine strings with R expressions (coerced to `as.character()`). `str_glue()` is designed to handle this more elegantly by treating text in curly braces `{}` as R code.

```{r}
str_concat(
  "hello", 
  10 * 10,
  "worlds",
  separator = " "
)

str_glue("hello {10*10} worlds")
```

Concatenate the elements of a character vector into a single string with `chr_collapse()`

```{r}
chr_collapse(c("abc", "def"))
chr_collapse(1:5, separator = ", ")
```

### Pattern matching operations

There are functions to **detect**, **locate**, **extract**, **replace**, **remove** and **split** patterns of characters in strings. Except for the detect family, suffixes `_first()`, `_nth()` and `last()` are used to specify pattern occurences within strings, and `_all()` to work with every occurrence.

A pattern by default is a regular expression, a type of string with a special set of characters defined that allow it to represent many different strings. If you wish to use a string literally as a position you can supply argument `fixed = TRUE`.

To get started we'll just work with a simple pattern

```{r}
strings <- c(
  "in the middle of the day, I eat lunch",
  "today's the day, forget yesterday",
  "only weeks to be found here",
  "days like this make me love Thursdays"
)

# Define a regular expression pattern for words containing "day"
pattern <- "\\w*day\\w*"
```

#### Detect

```{r}
# Does the string contain the pattern?
str_detect(strings, pattern)
str_detect_starts_with(strings, pattern)
str_detect_ends_with(strings, pattern)
```

#### Extract

```{r}
# Pull out the matching words from each string
str_extract_first(strings, pattern)
str_extract_nth(strings, pattern, 2)
str_extract_last(strings, pattern)
str_extract_all(strings, pattern)
```

#### Replace and Remove
```{r}
str_remove_first(strings, pattern)

str_replace_all(strings, pattern, "night")
```

#### Split
``` {r}
# Let's just split on spaces to get the words
str_split_all(strings, " ")
```

### Character vector organisation

Vectors are ordered, so you might like to reorder them.

```{r}
chr_sort(c("cherry", "apple", "date", "banana"))
```
